---
title: "Boston SER Talk code"
output: html_notebook
---
R Implementation -
Introduction to Parametric & Semi-parametric Estimators for Causal Inference
Laura Balzer & Jennifer Ahern September 28, 2018


During the workshop, we will complete this R exercise together. After the workshop, an answer key will be distributed.


Make sure libraries are loaded
```{r LOAD Libraries}
library(SuperLearner)
library(ltmle)


```

Load the data set CausalWorkshop.csv and assign it to object data.
```{r DATA IMPORT}
workshop = read.csv("/Volumes/GoogleDrive/My Drive/Edu/Epi PhD/Epidemiology/Talks Conferences and Seminars/SER/Boston 2018/DAT/CausalWorkshop.csv")
```

```{r Examine data}
#### Shows #Rows   #Col  ####
dim(data)
```

## Data Dictionary 
The observed data consist of the following variables
- W 1: Binary baseline covariate (values 0 and 1)
- W 2: Continuous baseline covariate (range from 0 to 1) - W 3: Continuous baseline covariate (centered at 0)
- W 4: Continuous baseline covariate (centered at 0)
- A: The exposure (1 for exposed; 0 for unexposed)
- Y : The outcome (range from 0 to 1).

```{r Setting the seed}
set.seed(1)
```

```{r STEP 1 -- Estimate the conditional mean outcome with parametric regression}
## Parametric G-computation
outcome.regression<- glm(Y~A+W1+W2+W3+W4, family='binomial', data=workshop)

summary(outcome.regression)
```

### Next, we obtain the expected (predicted) outcome for each observation under the exposure and given the covariates Eˆ(Y|A = 1,W) as well as the expected (predicted) outcome for each observation under no exposure and given the covariates Eˆ (Y |A = 0, W ).

(a) Copy the data set data into two new data frames data.txt and data.untxt. Then set all units in data.txt to be treated/exposed (A=1) and all units in data.untxt to be untreated/unexposed (A=0). The columns of a data frame can be accessed with the $ operator.

```{r Fitted regression to obtain predicted outcomes}
# copying the observed data
data.txt<- data.untxt <- workshop
# set A=1 in the data.txt and A=0 in data.untxt
data.txt$A <-1
data.untxt$A <- 0
# Verify with the colMeans function (which takes the mean in each data column)
colMeans(data.txt)
colMeans(data.untxt)
```

(b) Use the predict function to get the expected outcome for each observation i given the exposure and its covariates Eˆ(Yi|Ai = 1, Wi).

```{r}
predict.outcome.txt = predict(outcome.regression, newdata=data.txt, type='response')

```


